---
title: "Neuro-muscular analysis of FAMe study"
author: "Danielle Hiam"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 2
    code_folding: "hide"
  html_notebook:
    theme: lumen
    toc: yes
editor_options: 
  chunk_output_type: console
---
### Packages Required
```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r Packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(ggpubr)
library(gridExtra)
library(gtsummary)
library(lme4)
library(MASS)
library(kableExtra)
library(caret)
library(gratia)
library(mgcv)
library(betareg)
library(stargazer)
library(broom)
library(mice)
library(randomForest)
library(reshape2)
library(pheatmap)
library(car)
library(pdp)
library(corrplot)
library(FactoMineR)
library(missMDA)
library(factoextra)
library(RColorBrewer)
library(ggrepel)
library(PCAtools)
library(gt)
library(VIM)
library(skimr)
library(missForest)
```

```{r Data cleaning, eval=FALSE}
setwd("PATHWAY")

dat1= read_excel("FAMe_Data_v1.xlsx",
                 sheet = 1,  
                 na = c("","NA", "?", "#VALUE!", "Undetermined"),
                 skip = 1,
                 trim_ws = TRUE)
colnames(dat1)

#Subset data needed for analysis 
dat_sub = dat1 %>%
  dplyr::select("ID", "Phase","Contraceptive","Men_Phase_NM", "Menopause_Treatment", "Urine_E2", "Urine_Prog","IPAQ_NUM", "T", "Prog", "E2", "SHBG", "FEI", "FAI", "T_E2_Ratio", "MVPA/day_mins", "Protein (g/kg/day)","Age", "Weight","Height","BMI", "MuA_cm2", "Lean_MuA_cm2", "IntraFAT_cm2", "SubCutFAT_cm2","bonedensity", "Lean_kg", "Leg_lean_mass_kg", "Arm_Lean_Mass_kg", "ALM_kg", "ALMI_kgm2", "ALM_BMI_ratio", "Fat_kg", "Fat_percent", "AG_Ratio", "BMC_kg", "bonedensity", "Z_score", "Quad_CSA", "Quad_IMAT","E_1RM_kg", "MVC", "RTD", "PT_100", "PT_10", "PT_ratio", "VA", "HMAX_VL", "HMAX_VM", "HMAX_RF", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL",	"MDUR_VM", "MDUR_RF", "RMS_VL", "RMS_VM", "RMS_RF")%>%
  rename(MVPA=`MVPA/day_mins`) %>% 
  rename(Protein= `Protein (g/kg/day)`)%>%
  drop_na(MVC)  %>%
  mutate(across(Phase, as.factor))%>%
  mutate(Phase=fct_relevel(Phase, c("PRE-M", "PERI", "POST-M")))%>%
  mutate(Men_Phase_NM=fct_relevel(Men_Phase_NM, c("Phase 1","Phase 4", "NA (Phase 1)")))%>%
  mutate(across(IPAQ_NUM, as.factor))%>%
  mutate(across(Menopause_Treatment, as.factor)) %>%
  mutate(across(Contraceptive, as.factor)) 
glimpse(dat_sub)

dec=function(x, na.rm= F) (x/100)
dat_sub2 =dat_sub %>% 
  dplyr::mutate_at(c("VA", "HMAX_VL", "HMAX_VM", "HMAX_RF", "RMS_VL", "RMS_VM", "RMS_RF"),dec) 
#Remove outlier (physiologically impossible from MVPA)
dat_sub2[11, "MVPA"] <- NA
```

# Imputation 
```{r Missingness Patterns}
predictors= dat_sub2 %>%
  dplyr::select("Age", "MVPA", "Weight","Height", "Protein", "Quad_CSA", "Quad_IMAT", "bonedensity") 
Overview=skimr::skim(predictors)
pMiss <- function(x){sum(is.na(x))/length(x)*100}
ID_miss=as.data.frame(apply(predictors,2,pMiss)) #features (columns)
pred_miss=as.data.frame(apply(predictors,1,pMiss)) # rows
aggr_plot <- aggr(predictors, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(predictors), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

k=sqrt(88) 
M_KNN = kNN(dat_sub2,
            variable=c("Age", "MVPA", "Weight","Height", "Protein", "Quad_CSA", "Quad_IMAT", "bonedensity"),
              k=k)

Overview2=skimr::skim(M_KNN)
print(Overview2)
```

## Menstrual Phase
```{r Menstrual Phase,  message=FALSE, warning=FALSE, fig.width = 10, fig.height = 20}
PREM <- dat_sub2 %>%
  filter(Phase == "PRE-M") %>%
    drop_na(Men_Phase_NM)
summary(PREM)

vars= PREM%>%
  dplyr::select("E_1RM_kg", "MVC", "RTD", "PT_100", "PT_10", "PT_ratio", "VA", "HMAX_VL", "HMAX_VM", "HMAX_RF", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL",	"MDUR_VM", "MDUR_RF", "RMS_VL", "RMS_VM", "RMS_RF")
vars <- names(vars)

models = lapply(vars, function(x) {
  lm(substitute(i ~ Men_Phase_NM, list(i = as.name(x))), data = PREM)
})

y=lapply(models, summary)
names(y) <- vars # To add names of the model instead of list number
df=purrr::map_df(y, broom::tidy, .id = 'vars')%>%
  filter(term == "Men_Phase_NMPhase 4" | term != "(Intercept)")

knitr::kable(df) %>%
  kable_styling() %>%
  row_spec(which(df$term == "Men_Phase_NMPhase 4"  & df$p.value < 0.05), bold = T)
#write.csv(df, "results/csv/LM_MenstrualPhase.csv")

#Figures
plot_list<- vector('list', 21)
names(plot_list) <- vars
for(i in vars){
  plot_list[[i]] <- 
    ggplot(PREM, aes(x=Men_Phase_NM, y=.data[[i]], color=Men_Phase_NM)) +
    geom_point(size=3)+
    geom_boxplot()+
    theme(text = element_text(face="bold"),#, size=16), 
          panel.background = element_blank(), axis.line=  element_line(colour = "black"),
          legend.position = "none")
}
grid.arrange(grobs = plot_list, ncol = 4)
#ggsave(filename = "figures/Men_Phase_NM.pdf",  plot = marrangeGrob(plot_list, nrow = 4, ncol=4),    width = 25, height = 20)
```
## HRT
```{r HRT, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 20}
POST<- dat_sub2 %>%
  filter(Phase != "PRE-M")%>%
  mutate(Menopause_Treatment=fct_relevel(Menopause_Treatment, c("No", "Yes")))

vars= POST%>%
  dplyr::select("E_1RM_kg","MVC", "RTD", "PT_100", "PT_10", "PT_ratio", "VA", "HMAX_VL", "HMAX_VM", "HMAX_RF", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL",	"MDUR_VM", "MDUR_RF", "RMS_VL", "RMS_VM", "RMS_RF")
vars <- names(vars)

models = lapply(vars, function(x) {
  lm(substitute(i ~ Menopause_Treatment, list(i = as.name(x))), data = POST)
})
y=lapply(models, summary)
names(y) <- vars 
df=purrr::map_df(y, broom::tidy, .id = 'vars')%>%
  filter(term == "Menopause_TreatmentYes" | term != "(Intercept)")

knitr::kable(df) %>%
 kable_styling() %>%
 row_spec(which(df$term == "Menopause_TreatmentYes" & df$p.value < 0.05), bold = T)
#write.csv(df, "results/csv/LM_HRT.csv")

# Association with MenPhase
plot_list<- vector('list', 21)
names(plot_list) <- vars
for(i in vars){
  plot_list[[i]] <- 
    ggplot(POST, aes(x=Menopause_Treatment, y=.data[[i]], color=Menopause_Treatment)) +
    geom_point(size=3)+
    geom_boxplot()+
       theme(text = element_text(face="bold"),#, size=16), 
          panel.background = element_blank(), axis.line=  element_line(colour = "black"),
          legend.position = "none")
}
grid.arrange(grobs = plot_list, ncol = 4)
#ggsave(filename = "figures/Menopause_Treatment.pdf", plot = marrangeGrob(plot_list, nrow=4, ncol=4), width = 25, height = 20)
```
## Contraceptive Measure
```{r Contraception models, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 20}
PREM <- dat_sub2 %>%
  filter(Phase == "PRE-M")%>%
  mutate(Contraceptive=fct_relevel(Contraceptive, c("None", "OCP", "IUD", "Implant")))

vars= PREM%>%
  dplyr::select("E_1RM_kg","MVC", "RTD", "PT_100", "PT_10", "PT_ratio", "VA", "HMAX_VL", "HMAX_VM", "HMAX_RF", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL",	"MDUR_VM", "MDUR_RF", "RMS_VL", "RMS_VM", "RMS_RF")
vars <- names(vars)
models = lapply(vars, function(x) {
  lm(substitute(i ~ Contraceptive, list(i = as.name(x))), data = PREM)
})

y=lapply(models, summary)
names(y) <- vars 
df=purrr::map_df(y, broom::tidy, .id = 'vars')%>%
  filter(term != "(Intercept)")
knitr::kable(df) %>%
 kable_styling() %>%
row_spec(which(df$p.value < 0.05), bold = T)
write.csv(df, "results/csv/LM_Contraceptive method.csv")

# Association with MenPhase
plot_list<- vector('list', 21)
names(plot_list) <- vars
for(i in vars){
  plot_list[[i]] <- 
    ggplot(PREM, aes(x=Contraceptive, y=.data[[i]], color=Contraceptive)) +
    geom_point(size=3)+
    geom_boxplot()+
       theme(text = element_text(face="bold"),#, size=16), 
          panel.background = element_blank(), axis.line=  element_line(colour = "black"),
          legend.position = "none")
}
grid.arrange(grobs = plot_list, ncol = 4)
#ggsave(filename = "figures/Contraceptive.pdf", plot = marrangeGrob(plot_list, nrow=4, ncol=4), width = 25, height = 20)
```

# Association of Neuro-muscular variables with age
```{r Model Fit, fig.show='hide', message=FALSE, warning=FALSE}
vars= dat_sub2%>%
  dplyr::select("E_1RM_kg","MVC", "RTD", "PT_100", "PT_10", "PT_ratio", "VA", "HMAX_VL", "HMAX_VM", "HMAX_RF", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL",	"MDUR_VM", "MDUR_RF", "RMS_VL", "RMS_VM", "RMS_RF", "Quad_CSA", "Quad_IMAT")
vars <- names(vars)

plot_list<- vector('list', 21)
names(plot_list) <- vars
for(i in vars){
  plot_list[[i]] <- 
    ggplot(dat_sub2, aes(x=MVPA, y=.data[[i]])) + ### CHANGE X to protein, MVPA manually
    geom_point(size=3)+
    geom_smooth(method = "lm")+
    scale_color_gradientn(colours = rainbow(5))+
    theme(text = element_text(face="bold", size=16), panel.background = element_blank(), axis.line=  element_line(colour = "black"))
}
#ggsave(filename = "figures/MVPA_plots_LM.pdf", plot = marrangeGrob(plot_list, nrow=3, ncol=3), width = 20, height = 20)

# Run models for residual 
models = lapply(vars, function(x) {
  lm(substitute(i ~ Age + MVPA + Protein, list(i = as.name(x))), data = dat_sub2)
})
y=lapply(models, summary)

```

# Which Model?
```{r Modelling, message=FALSE, warning=FALSE, fig.width = 15, fig.height = 20}
# Run gam or lm models 
vars = dat_sub2 %>%
  dplyr::select("E_1RM_kg", "RTD", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL", "MDUR_VM", "MDUR_RF", "MVC", "PT_100", "PT_10", "PT_ratio", "Quad_CSA", "Quad_IMAT")
vars <- names(vars)

vars2 = dat_sub2 %>%
  dplyr::select("E_1RM_kg", "MVC")
vars2 <- names(vars2)

# Fit models for comparison: full model (with MVPA and Protein) vs null model (only Age)
models_comparison <- list()
for (var in vars2) { # change vars vars 2 depending model being run. 
    model_full <- gam(as.formula(paste(var, "~ s(Age) + MVPA + Protein ")), 
                    method = 'REML', 
                    data = dat_sub2)
    model_lm_full <- gam(as.formula(paste(var, "~ Age+ MVPA + Protein + Quad_CSA")), 
                    method = 'REML', 
                    data = dat_sub2)
    model_null <- gam(as.formula(paste(var, "~ s(Age) + Quad_CSA")), 
                    method = 'REML', 
                    data = dat_sub2)
    model_lm_null <- gam(as.formula(paste(var, "~ Age + Quad_CSA")), 
                    method = 'REML', 
                    data = dat_sub2)
    
    aic_full_gam <- AIC(model_full)
    aic_null_gam <- AIC(model_null)
    aic_full_lm <- AIC(model_lm_full)
    aic_null_lm <- AIC(model_lm_null)
    aic_comparison <- data.frame(
      model = c("Full GAM", "Null GAM", "Full LM", "Null LM"),
      AIC = c(aic_full_gam, aic_null_gam, aic_full_lm, aic_null_lm)
    )
    best_model <- aic_comparison[which.min(aic_comparison$AIC), ]
    models_comparison[[var]] <- data.frame(
      variable = var,
      aic_full_gam = aic_full_gam,
      aic_null_gam = aic_null_gam,
      aic_full_lm = aic_full_lm,
      aic_null_lm = aic_null_lm,
      best_model_name = best_model$model,
      best_model_aic = best_model$AIC
    )
}

comparison_df <- do.call(rbind, models_comparison)
print(comparison_df)
#Essentially all models are better with MVPA and Protein

# MULTICOLLINEARITY
vars = dat_sub2 %>%
  dplyr::select("E_1RM_kg", "RTD", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL", "MDUR_VM", "MDUR_RF", "MVC", "PT_100", "PT_10", "PT_ratio", "Quad_CSA")
vars <- names(vars)
vars2 = dat_sub2 %>%
  dplyr::select("E_1RM_kg", "MVC")
vars2 <- names(vars2)

models_comparison <- list()
for (var in vars2) {
    # Fit the model with Age, MVPA, Protein, and Quad_CSA as predictors
    model <- lm(as.formula(paste(var, "~ Age + MVPA + Protein + Quad_CSA ")),
                 data = dat_sub2)
    vif_values <- vif(model)  
    high_vif_vars <- names(vif_values)[vif_values > 5] 
    models_comparison[[var]] <- list(
      model = model,
      vif_values = vif_values,
      high_vif_vars = high_vif_vars )
}

vif_results <- lapply(models_comparison, function(x) {
  data.frame(variable = names(x$vif_values), 
             vif = x$vif_values, 
             high_vif = ifelse(x$vif_values > 5, "Yes", "No"))
})
vif_df <- do.call(rbind, vif_results)
#VIF scores were calculated with Age as a linear term, as it is tricky to calc when age is a smooth. 
# Mutlicollinearity is not an issue
```

# Run GAM models
```{r}
set.seed(123)
vars_gam = dat_sub2 %>%
  dplyr::select("E_1RM_kg", "RTD", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL", "MDUR_VM", "MDUR_RF", "MVC", "PT_100", "PT_10", "PT_ratio", "Quad_CSA", "Quad_IMAT")
vars_gam <- names(vars_gam)
k_check_results_df <- data.frame(Variable = character(), 
                                 `s.Age.` = numeric(), 
                                 edf = numeric(), `k-index` = numeric(), 
                                 `p-value` = numeric(), 
                                 stringsAsFactors = FALSE)

# Fit GAM models 
models = lapply(vars_gam, function(x) {
  model <- gam(substitute(i ~ s(Age)+ MVPA + Protein , list(i = as.name(x))), 
               method = 'REML', 
               data = dat_sub2)
  return(model)
})
names(models) <- vars_gam 
print(names(models))

k_check=lapply(models, k.check)
for (var in names(k_check)) {
  result <- k_check[[var]]
  result_row <- data.frame(
    Variable = var,
    `k'` = result[1],     
    edf = result[2],          
    `k-index` = result[3],    
    `p-value` = result[4],    
    stringsAsFactors = FALSE  
  )
  k_check_results_df <- rbind(k_check_results_df, result_row)
}

print(k_check_results_df)

pdf('figures/Residuals/Gam_residuals.pdf')
names(models) <- vars_gam 
par(mfrow=c(2,2)) 
invisible(lapply(names(models), function(x) {
  gam.check(models[[x]])  
  title(main = x)         
}))
dev.off()

m=gam(E_1RM_kg ~ s(Age)+ Protein + MVPA + Quad_CSA, data = dat_sub2)
summary(m)

## RUN GAM MODELS
y=models
names(y) <- vars_gam 
df=purrr::map_df(y, broom::tidy, .id = 'vars') 
knitr::kable(df) %>%
  kable_styling() %>%
  row_spec(which(df$term == "s(Age)"  & df$p.value < 0.05), bold = T)
write.csv(df, "results/csv/gam models summary_impute.csv")


### MVC 1RM with Quad CSA
vars2 = dat_sub2 %>%
  dplyr::select("E_1RM_kg", "MVC")
vars2 <- names(vars2)
k_check_results_df <- data.frame(Variable = character(), 
                                 `s.Age.` = numeric(), 
                                 edf = numeric(), `k-index` = numeric(), 
                                 `p-value` = numeric(), 
                                 stringsAsFactors = FALSE)

# Fit GAM models 
models = lapply(vars2, function(x) {
  model <- gam(substitute(i ~ s(Age) + MVPA + Protein + Quad_CSA , list(i = as.name(x))), 
               method = 'REML', 
               data = dat_sub2)
  return(model)
})
names(models) <- vars2 # To add names of the model instead of list number
print(names(models))

k_check=lapply(models, k.check)
for (var in names(k_check)) {
  result <- k_check[[var]]
  result_row <- data.frame(
    Variable = var,
    `k'` = result[1],     
    edf = result[2],          
    `k-index` = result[3],    
    `p-value` = result[4],    
    stringsAsFactors = FALSE  
  )
  k_check_results_df <- rbind(k_check_results_df, result_row)
}
print(k_check_results_df)

# GAM MODELS with adjusted CSA
y=models
names(y) <- vars2
df=purrr::map_df(y, broom::tidy, .id = 'vars') 
knitr::kable(df) %>%
  kable_styling() %>%
  row_spec(which(df$term == "s(Age)"  & df$p.value < 0.05), bold = T)

```

# Exploratory Analysis of Predictor Variables
# **Identify trends and changes in the data**
**Steep Indices vs. Change Points**\
-Steep Indices: Refers to the indices in the data where the first derivative of the GAM model falls below a specific threshold (75% of the mean first derivative).
Indicates where the response (RTD) experiences a significant decline, suggesting a point where aging may affect RTD more drastically.\
-Change Points:Refers to the indices where the sign of the second derivative changes.
Indicates points where the acceleration or deceleration of the response changes, suggesting a shift in the trend (e.g., transitioning from increasing RTD to decreasing RTD).

```{r Steepness and changes indices}
source("C:/Users/dhiam/OneDrive - Deakin University/Manuscripts/NeuroMusc FAMe/Analysis/derivative_functions.R")
# Fit the GAM model
mod <- gam("Change Var" ~ s(Age) + MVPA + Protein, method = 'REML', data = dat_sub2, na.action = na.exclude)

# Calculate derivatives
fd <- derivSimulCI_adjusted(mod)
 first_derivative <- fd$Age$deriv
 second_derivative <- fd$second_deriv
 evaluated_ages <- fd$eval
 mean_derivative <- mean(first_derivative, na.rm = TRUE)
 threshold <- mean_derivative * 0.75  
# Find indices where the first derivative is below the threshold
 steep_indices <- which(first_derivative < threshold)
 second_derivative[is.na(second_derivative)] <- 0  
# Find change points using the second derivative
 change_points <- which(diff(sign(second_derivative)) != 0)
 
# Extract the corresponding ages for steepness indices and change points
steep_ages <- evaluated_ages[steep_indices]
change_ages <- evaluated_ages[change_points]
 cat("Steepness Indices (First Derivative):\n")
print(steep_ages)

cat("\nChange from Increase to Decrease (Second Derivative):\n")
print(change_ages)
 
new_data <- data.frame(Age = evaluated_ages,
                       MVPA = mean(dat_sub2$MVPA, na.rm = TRUE),
                       Protein = mean(dat_sub2$Protein, na.rm = TRUE))  

 # Predicted RTD for the evaluated ages
predicted_RTD <- predict(mod, newdata = new_data)
steep_ages <- evaluated_ages[steep_indices]
steep_points_df <- data.frame(Start_Age = steep_ages)
steep_points_df <- steep_points_df[!duplicated(steep_points_df$Start_Age), ]

results_df <- data.frame(
   Age = evaluated_ages,
   RTD = predicted_RTD,
   First_Derivative = first_derivative,
   Second_Derivative = second_derivative)

Steep_change_df <- results_df %>%
  mutate(
    Steep_Age = ifelse(Age %in% evaluated_ages[steep_indices], Age, NA),
    Change_Age = ifelse(Age %in% evaluated_ages[change_points], Age, NA)
  )

Steep_change_df <- Steep_change_df %>%
  filter(!is.na(Steep_Age) | !is.na(Change_Age))
 colnames(results_df)[colnames(results_df) == "Age.1"] <- "Second_Derivative"

  ggplot(results_df, aes(x = Age)) +
   geom_line(aes(y = First_Derivative), color = 'blue', size = 1, linetype = "dashed") +
   geom_line(aes(y = Second_Derivative), color = 'red', size = 1, linetype = "dotted") +
   geom_hline(yintercept = threshold, linetype = "dotted", color = "black") +
   geom_vline(xintercept = evaluated_ages[steep_indices], color = "green", linetype = "dashed") +
   geom_vline(xintercept = evaluated_ages[change_points], color = "purple", linetype = "dashed") +
   labs(title = "GAM Fit and Derivatives",
        y = "Value",
        x = "Age") +
   theme_minimal()
```

# PCA analysis
```{r PCA data cleaning for all variables, eval=FALSE}
#Subset data needed for analysis 
dat_sub3 = dat_sub2 %>%
  dplyr::select("ID", "Phase","MVPA", "Age", "Protein","Weight","Height", "Lean_MuA_cm2", "IntraFAT_cm2", "Quad_CSA","Quad_IMAT","Lean_kg", "Fat_kg",  "bonedensity",  "E_1RM_kg", "MVC", "RTD", "PT_100", "PT_10", "PT_ratio", "VA", "HMAX_VL", "HMAX_VM", "HMAX_RF", "MAMP_VL", "MAMP_VM", "MAMP_RF", "MDUR_VL",	"MDUR_VM", "MDUR_RF", "RMS_VL", "RMS_VM", "RMS_RF","T", "E2", "Prog", "SHBG", "FEI", "FAI", "T_E2_Ratio") %>%
  drop_na(MVC) %>%
    filter(Phase == "POST-M")
#Impute missing data for predictors
set.seed(123)
library(VIM)
# IMPUTATION
predictors= dat_sub3 %>%
  dplyr::select("Age", "MVPA", "Weight","Height", "Protein", "Quad_CSA", "Quad_IMAT", "bonedensity", "T", "E2", "Prog", "SHBG", "FEI", "FAI", "T_E2_Ratio") 
Overview=skimr::skim(predictors)
pMiss <- function(x){sum(is.na(x))/length(x)*100}
ID_miss=as.data.frame(apply(predictors,2,pMiss)) #features (columns)
pred_miss=as.data.frame(apply(predictors,1,pMiss)) # rows
aggr_plot <- aggr(predictors, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(predictors), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
k=sqrt(42) #sample size
M_KNN = kNN(dat_sub3,
            variable=c("Age", "MVPA", "Weight","Height", "Protein", "Quad_CSA", "Quad_IMAT", "bonedensity", "T", "E2", "Prog", "SHBG", "FEI", "FAI", "T_E2_Ratio"),
              k=k)

Overview2=skimr::skim(M_KNN)
print(Overview2)

corrPlot <- column_to_rownames(final_df, var="Dim.1.Variable")

name_mapping <- c(
  "Age" = "Age",  
  "MVPA" = "MVPA",  
  "Protein" = "Protein",  
  "bonedensity" = "BMD",  
  "Quad_CSA" = "Quadriceps lean CSA",  
  "Quad_IMAT" = "Quadriceps IntraFat CSA",  
  "E2" = "Oestradiol",  
  "FEI" = "Free Oestrogen Index", 
  "Prog" = "Progesterone", 
  "T" = "Testosterone",
  "FAI" = "Free Testosterone Index" 
)

vars_original <- c("Age", "MVPA", "Protein", "BMD", "Quadriceps lean CSA", 
                   "Quadriceps IntraFat CSA", "Oestradiol", "Free Oestrogen Index", 
                   "Progesterone", "Testosterone", "Free Testosterone Index")

vars_mapped <- c("Age", "MVPA", "Protein", "bonedensity", "Quad_CSA", 
                 "Quad_IMAT", "E2", "FEI", "Prog", "T", "FAI")
corrPlot <- corrPlot[match(vars_mapped, rownames(corrPlot)), ]
corr <- corrPlot %>%
  dplyr::select(Dim.1.Correlation, Dim.2.Correlation, Dim.3.Correlation, Dim.4.Correlation)

p <- corrPlot %>%
  dplyr::select(Dim.1.P_value, Dim.2.P_value, Dim.3.P_value, Dim.4.P_value)

corr <- as.matrix(corr)
p <- as.matrix(p)

rownames(corr) <- vars_original  
colnames(corr) <- c("PC1", "PC 2", "PC3", "PC4")  
rownames(p) <- vars_original
colnames(p) <- c("PC1", "PC 2", "PC3", "PC4")


wide_data <- corrPlot2 %>%
  pivot_wider(
    names_from = PC,
    values_from = c(Correlation, P.Value),
    names_glue = "{PC}.{.value}"
  )
vars_mapped <- c("MVC", "E_1RM_kg","RTD","PT_10", "PT_100")
corrPlot2 <- column_to_rownames(wide_data, var="Outcome")
corrPlot2 <- corrPlot2[match(vars_mapped, rownames(corrPlot2)), ]

corr2 <- corrPlot2 %>%
  dplyr::select(Dim.1.Correlation, Dim.2.Correlation, Dim.3.Correlation, Dim.4.Correlation)
p2 <- corrPlot2 %>%
  dplyr::select(Dim.1.P.Value, Dim.2.P.Value, Dim.3.P.Value, Dim.4.P.Value)
cols <- c("PC1", "PC2", "PC3", "PC4")
colnames(corr2) <- cols
colnames(p2) <- cols
corr2<-as.matrix(corr2)
p2<-as.matrix(p2)
```

# PCA Figures
```{r}

tiff("figures/corrplots_combined.tiff", width = 20, height = 10, units = "in", res = 600)
par(mfrow = c(1, 2),
    mar = c(6, 10, 6, 6),  # Standard margins
    oma = c(0,0,0,0),   # Increased top outer margin for alignment
    pty = "s")

c1<-corrplot(corr, 
             insig = 'label_sig',
             order = "original",
             p.mat = p,
             sig.level = 0.05,
             method = "color", 
             type = "full", 
             tl.cex = 1.8,           
             tl.col = "white",       
             tl.srt = 45,
             col = COL2('PRGn'),
             cl.pos = 'r',
             addgrid.col = "gray",
             cl.cex = 1.5,         
             cl.ratio = 0.5, 
            mar = c(6, 6, 8, 0))  

text(1:4, 12.1, 
     expression("PC 1", "PC 2", "PC 3", "PC 4"), 
     cex = 1.8, font = 2,
     srt =45)

mtext("2A: Correlation of predictors with PCs 1-4", 
      side = 3, line = 1.5, cex = 1.5, font = 2)


mtext(side = 2, line = -18, at = 11:1, 
      text = c("Age", "MVPA", "Protein intake", "BMD", "Quadriceps Lean CSA", "Quadriceps Intramuscular fat CSA", "Oestradiol", "Free Oestradiol Index", "Progesterone", "Testosterone", "Free Testosterone Index"),
      cex = 1.5, font = 1, adj = 1, las = 2)


c2<-corrplot(corr2,
             order = "original",
             insig = 'label_sig',
             p.mat = p2,
             sig.level = 0.05,
             method = "color", 
             type = "full", 
             tl.cex = 1.8,           
             tl.col = "white",       
             tl.srt = 45,
             pch.cex = 3,
             col = COL2('PRGn'),
             cl.pos = 'r',
             addgrid.col = "gray",
             cl.cex = 1.5,        
             cl.ratio = 0.3,      
          mar = c(6, 5, 12, 2))   

text(1:4, 5.9, 
     expression("PC 1", "PC 2", "PC 3", "PC 4"), 
     cex = 1.8, font = 2,
     srt =45, adj = c(0.5, 0))

mtext("2B: Correlation of neuromuscular measures with PCs 1-4", 
      side = 3, line = 1.5, cex = 1.5, font = 2)

mtext(side = 2, line =-7, at = 5:1,  
      text = c("MVC (N\U00B7m)", 
               "e1RM (kg)", 
               expression(paste("RTD (N\U00B7m.s"^-1, ")")),
               expression(PT[10]~"(N\U00B7m)"), 
               expression(PT[100]~"(N\U00B7m)")),
      cex = 1.5, font = 2, adj = 1, las = 2)
dev.off()
```

# Quad_IMAT
```{r Quad_IMAT}
model<- gam(Quad_IMAT ~ s(Age)+ MVPA + Protein , 
               method = 'REML', 
               data = dat_sub2)

summary(model)
k.check(model)
```


